Технология использования открытых данных Федеральной налоговой службы России для экономико-географического анализа

### 2024-11-18 fts-open-data-in-economic-geography/Paper.Rmd
Update abstract (once more), #5

***Аннотация.*** В статье представлена технология создания геокодированного табличного набора данных о субъектах малого и среднего предпринимательства (МСП) в России на основе трёх наборов открытых данных Федеральной налоговой службы (ФНС России). Получаемый в результате использования технологии набор данных охватывает всю территорию России и содержит информацию, детализированную до уровня конкретных юридических лиц и индивидуальных предпринимателей. Таблица включает следующие поля (колонки): ИНН, ОГРН, вид (юридическое лицо, индивидуальный предприниматель, глава крестьянского (фермерского) хозяйства), категория (микропредприятие, малое, среднее предприятие), имя или наименование, место жительства или местонахождение (субъект федерации, район, город, населённый пункт), код основного вида деятельности по ОКВЭД, доходы, расходы, среднесписочная численность работников. Данные о доходах, расходах и численности работников доступны за каждый год с 2018 г., остальные — за каждый месяц с августа 2016 г. Статья излагает воспроизводимую методику обработки исходных данных и демонстрирует её применение для получения и разведочного анализа сведений о фирмах в сфере сельского, лесного хозяйства и рыболовства. Дополнением к статье является открытое консольное приложение на языке программирования Python, реализующее описанную технологию. Показано, что открытые данные ФНС России с помощью предложенного алгоритма можно использовать для решения ряда аналитических и исследовательских задач в области экономической географии, в особенности тех, где требуются дезагрегированные сведения или пространственная детализация до населённых пунктов. Благодаря географическим координатам получаемый набор данных может без дополнительной обработки использоваться для картографирования. Наличие кодов ОКТМО позволяет объединять его с официальными статистическими сведениями.

@@

***Abstract.*** The paper presents a methodology for the creation of a geocoded tabular dataset of small and medium-sized enterprises (SMEs) in Russia based on three open datasets provided by the Federal Tax Service (FTS of Russia). The resulting dataset encompasses the entire territory of Russia. The information is provided at the level of specific legal entities or individual entrepreneurs. The dataset is structured as a CSV file comprising the following fields: tax number, registration number, legal status (legal entity, individual entrepreneur, head of farming company), SME category (microbusiness, small or medium-sized business), name, address (region, district, city, settlement), main activity code according to NACE, income, expenses, and average list number of employees. The dataset includes revenue, expenses, and employee data from 2018 onwards, with yearly granularity, and all other variables from August 2016 onwards, with monthly granularity. The article outlines a reproducible methodology for raw data processing and demonstrates its application to obtain a dataset of firms in the agriculture, forestry and fisheries sectors and conduct its exploratory data analysis. A reference implementation of the described technology is provided in a form of an open-source Python command-line tool. The paper shows that with the help of proposed technique, FTS open data can be used to solve various analytical and academic tasks in economical geography, in particular those where the disaggregated information or spatial resolution up to settlements is required. In addition, the incorporation of geographic coordinates into the dataset enables direct mapping without the need for additional processing. The presence of OKTMO codes allows for seamless integration with official statistical information.

>>>>

***Аннотация.*** В статье представлена технология создания геокодированного табличного набора данных о субъектах малого и среднего предпринимательства (МСП) в России на основе открытых данных Федеральной налоговой службы (ФНС России). Получаемый в результате применения технологии набор данных охватывает всю территорию России и содержит информацию, детализированную до уровня конкретных юридических лиц и индивидуальных предпринимателей. Таблица включает следующие поля (колонки): ИНН, ОГРН, вид (юридическое лицо, индивидуальный предприниматель, глава крестьянского (фермерского) хозяйства), категория (микропредприятие, малое, среднее предприятие), имя или наименование, место жительства или местонахождение (субъект федерации, район, город, населённый пункт), код основного вида деятельности по ОКВЭД, доходы, расходы, среднесписочная численность работников. Данные о доходах, расходах и численности работников доступны за каждый год с 2018 г., остальные — за каждый месяц с августа 2016 г. Статья излагает воспроизводимую методику обработки исходных данных и демонстрирует её применение для получения и разведочного анализа сведений о фирмах в сфере сельского, лесного хозяйства и рыболовства. Дополнением к статье является открытое консольное приложение на языке программирования Python, реализующее описанную технологию. Показано, что открытые данные ФНС России с помощью предложенного алгоритма можно использовать для решения ряда аналитических и исследовательских задач в области экономической географии, в особенности тех, где требуются дезагрегированные сведения или пространственная детализация до населённых пунктов. Благодаря географическим координатам получаемый набор данных может без дополнительной обработки использоваться для картографирования. Наличие кодов ОКТМО позволяет объединять его с официальными статистическими сведениями.

@@

***Abstract.*** This paper presents a methodology for the creation of a geocoded tabular dataset of small and medium-sized enterprises (SMEs) in Russia based on open data provided by the Federal Tax Service (FTS) of Russia. The resulting dataset encompasses the entire territory of the country. The data is provided at the level of individual SMEs. The dataset is structured as a CSV file comprising the following fields: tax number, registration number, legal status (juridical person, sole trader, head of peasant (farm) enterprise), SME category (microbusiness, small-sized business, medium-sized business), name, address (region, district, city, settlement), main activity code according to NACE, income, expenses, and average list number of employees. The dataset includes revenue, expenses, and employee data from 2018 onwards, with yearly granularity, and all other variables from August 2016 onwards, with monthly granularity. The article presents a reproducible methodology for the processing of raw FTS data and illustrates its application in the generation and exploratory data analysis of a dataset comprising firms in the agriculture, forestry and fisheries sectors. A reference implementation of the described technology is provided in the form of an open-source Python command-line tool. The paper demonstrates that the proposed technique enables the utilisation of FTS open data to address a range of analytical and academic tasks in the field of economic geography, particularly those benefiting from disaggregated information or requiring spatial resolution at the settlement level. Furthermore, the incorporation of geographic coordinates into the dataset facilitates direct mapping without the necessity for additional processing. The inclusion of municipal codes allows for the seamless integration with official statistical information.
